A Hitchhiker's Guide to Reproducible Research" # replace with something serious

### 2021-05-04 manuscript.Rmd
explain interplay of Docker and Make close #13

This command will go through all R Markdown files, and their YAML metadata, in your repository and infer a Makefile from them:

@@

@@

Modern data analysis is only feasible because we rely on much external software.
One downside is that a considerable number of external dependencies may hinder collaboration and long term reproducibility.
The critical problem is the fragility of a software setup.
Many components have to fit perfectly together, and each change, e.g. an update, risks that balance.
Updates are, of course, very much necessary.
Another problem is that the software environment is also variable and what works on one machine might not work on any other.
Docker circumvents such issues.
Instead of leaving it to the user which software should be installed how on their computer, Docker provides all software used in an analysis, down to the operating system.
Docker does this without interfering with the already installed software by using a virtual software environment that is independent of the host software environment.
Such a snapshot of the software stack is called "image".
Of course, packaging all needed software in such an image requires considerable amounts of storage.
However, two strategies help to keep the storage requirements manageable.
First, there is a community that maintains images for particular purposes.
For example, using R [rocker-citation].
Such images only contain what is necessary for a specific purpose.
Users can then install whatever they additionally need on top of these purpose-built images.
The image that was used for this article uses `r round(as.numeric(fs::file_size("reprotutorial.sif"))/1024^3, 2)`Gb of disc space.
It does include Ubuntu (a Linux operating system) R, RStudio, LaTeX as well as many R packages like the tidyverse (amounting to `r installed.packages() %>% as.data.frame() %>% filter(is.na(Priority)) %>% nrow()` packages.)
Second, instead of saving a binary image of the software, one can also save a textual description of the programmatic steps to install the software in a `Dockerfile`.
Such files are tiny (the `r link("https://github.com/aaronpeikert/repro-tutorial/blob/master/Dockerfile", "Dockerfile")` for this project is `r round(as.numeric(fs::file_size("Dockerfile"))/1024^1, 2)`Kb big), but they rely on the assumption that all software repositories are available and contain the correct software version.
The `repro`-package lets you specify required packages as part of the metadata of the RMarkdown:

@@

With the R version that you currently use, this specification is then translated into a `Dockerfile`.

@@

@@

One usecase in which Docker sometimes can not be used is in high performance computing (HPC).
The simulation we did is computationally expensive enough to warrant the use of a HPC cluster.
To run the simulation on a cluster we require the exact same software, not only on one or two computers, but potentially hundreds.
Docker is a natural solution for this problem and is the defacto standard for cloud computing.
However, Docker requires admin rights, while this is no issue on cloud computing infrastructure it is an issue in classical HPC environments.
Singularity was designed to run without admin rights, while maintaining full compatibility with Docker images.
If you already have a Docker image, you can choose if you want to run this image with Docker or with Singularity.
Anyway, from within the container i.e. for your analysis it does not make any difference.

>>>>

This command will go through all R Markdown files, and their YAML metadata, in your repository and infer a `Makefile` from them:

@@

The generated `Makefile` automatically contains targets for all R Markdown documents.
To, for example, generate the `README.md` from the `README.Rmd` you can run:

``` bash
# this should be run in the terminal
make README.md
```

It is common practice to have a target all, that encompasses all important files of the project as the first target.
If you do not specify a target, Make defaults to the first target, which is usually `all`.
`repro` reminds you of that and suggests which file you might want to add to the target `all`.
After you have done so, you can then run in the terminal:

``` bash
# this should be run in the terminal
make
```

And it will fully reproduce the project, if you have all required software installed.

@@

Modern data analysis is only feasible because we rely on much external software.
One downside is that a considerable number of external dependencies may hinder collaboration and long term reproducibility.
The critical problem is the fragility of a software setup.
Many components have to fit perfectly together, and each change, e.g. an update, risks that balance.
Updates are, of course, very much necessary.
Another problem is that the software environment is also variable and what works on one machine might not work on any other.
Docker circumvents such issues.
Instead of leaving it to the user which software should be installed how on their computer, Docker provides all software used in an analysis, down to the operating system.
Docker does this without interfering with the already installed software by using a virtual software environment that is independent of the host software environment.
Such a snapshot of the software stack is called "image".
Of course, packaging all needed software in such an image requires considerable amounts of storage.
However, two strategies help to keep the storage requirements manageable.
First, there is a community that maintains images for particular purposes.
For example, using R [rocker-citation].
Such images only contain what is necessary for a specific purpose.
Users can then install whatever they additionally need on top of these purpose-built images.
The image that was used for this article uses `r round(as.numeric(fs::file_size("reprotutorial.sif"))/1024^3, 2)`Gb of disc space.
It does include Ubuntu (a Linux operating system) R, RStudio, LaTeX as well as many R packages like the tidyverse (amounting to `r installed.packages() %>% as.data.frame() %>% filter(is.na(Priority)) %>% nrow()` packages.)
Second, instead of saving a binary image of the software, one can also save a textual description of the programmatic steps to install the software in a `Dockerfile`.
Such files are tiny (the `r link("https://github.com/aaronpeikert/repro-tutorial/blob/master/Dockerfile", "Dockerfile")` for this project is `r round(as.numeric(fs::file_size("Dockerfile"))/1024^1, 2)`Kb big), but they rely on the assumption that all software repositories are available and contain the correct software version.
For archiving, you therefore have to store the image, as explained in [Archiving Docker].
The `repro`-package lets you specify required packages as part of the metadata of the RMarkdown:

@@

With the R version that you currently use, this specification is then translated into a `Dockerfile`:

@@

From the `Dockerfile` you can then build an image.
While you can do so manually we prefer to use Make, see [Internal Dependencies].
The `Makefile` build by `repro` does contain a recipe for creating the `Dockerfile`.

``` bash
make docker
```

The `Makefile` from `repro` does also include a mechanism to send commands conditionally through the Docker container.
This mechanism is triggered by `DOCKER=TRUE` in the make call:

``` bash
# this should be run in the terminal
make DOCKER=TRUE
```

@@

One usecase in which Docker sometimes can not be used is in high performance computing (HPC).
The simulation we did is computationally expensive enough to warrant the use of a HPC cluster.
To run the simulation on a cluster we require the exact same software, not only on one or two computers, but potentially hundreds.
Docker is a natural solution for this problem and is the defacto standard for cloud computing.
However, Docker requires admin rights, while this is no issue on cloud computing infrastructure it is an issue in classical HPC environments.
Singularity was designed to run without admin rights, while maintaining full compatibility with Docker images.
If you already have a Docker image, you can choose if you want to run this image with Docker or with Singularity.
Anyway, from within the container i.e. for your analysis it does not make any difference.
Like with Docker, the default `Makefile` of `repro` has a mechanism to automatically send all commands to Singularity.
It is triggered by `SINGULARITY=TRUE`:

``` bash
make SINGULARITY=TRUE
```
